Assessing United States county-level exposure for tropical cyclone impact research

### 2020-07-21 ehp_revision/response_to_reviewers.Rmd

The second issue is exposure measurement error/misclassification. When the 
exposure (e.g., binary exposure or continuous measure of intensity of exposure)
is only measured at the county level, all the individuals in that county are 
assigned that same exposure. However, the hazards included in this study will 
vary across each county. For some hazards, this variation will be more stark. 
For example, tornadoes can be very local, causing extreme damage to a row of 
homes and none to the homes across the street. Some types of floods can also 
be very localized, rather than creating widespread flooding throughout 
the county. In both cases, when a county-level exposure assessment is made
based on an event in the county in association with the storm, there may be
many individuals in the county who did not experience that hazard at their 
residence. The other hazards we included---storm-associated winds and rainfall---will
also have some variation across counties, but this will often not be as stark. 
The rainfields for tropical cyclones are very large, and while there are rainbands
within the storm that might have particularly high rates of precipitation, these
progess over the course of the storm, and it is unlikely that a county will have
one area that experienced very extreme precipitation while another experienced
very little. Winds might have more variation, especially in coastal counties and 
counties near the storm's central track, since winds decay quickly as you move
from the storm's center, and storms typically weaken rapidly once they make 
landfall and can no longer draw on warm ocean waters for energy.

\begin{shaded}
\textbf{R1 Comment 2:}
Are there opportunities to expand this approach internationally? Some decisions
(e.g., county-level) data may restrict cross-country comparisons. Can the
authors discuss this? 
\end{shaded}

**Response:**

- Differs by exposure metric. Easier for some (wind, rain) than for others (flood, 
tornado)
- For rain, we use a re-analysis product only available for the continental US (right?).
However, there are similar gridded re-analysis products available for other regions 
worldwide that could be used in a similar way.
- For tornadoes, there may be national databases in other spots similar to the tornado
database that we use, but as far as we're aware, no international one. 
- Same for floods. 
- For wind, the wind model could be generalizable with some tweaks. These include: 
  + More landmasks
  + Direction of storm movement for Northern vs Southern Hemispheres
  + Different wind averaging periods for the recorded best tracks data in different 
  storm basins.
  + Our wind model was derived with US data and so may need bias-correction or a shift
  to a West Pacific--based model

> "Tornadoes long have been recognized as a global phenomenon (e.g., Wegener 1917; 
Feuerstein et al. 2005), having been recorded in every continent except Antarctica. 
As such, numerous nationas share an interest in improving assessment of their 
damage. International research and involvement in tornado survey work and damage
and intensity scales is well underway." [@edwards2013tornado]

@@

We have added some discussion of this point...

- Future research could expand the work by [@grabich2015measuring] to determine
the extent of these issues with use of FEMA disaster declarations for
epidemiological exposure assessment. This would need to go beyond the comparison
with exposure assessment based on storm hazards, because sometimes the
characteristics of the affected community modifies the amount of disaster that
results from hazards of a certain severity. For example, a community that it
exposed often might be more "hardened" against damage from the storm (any work
from Seth on this) and so might experience less infrastructure damage from a
tropical cyclone with certain characteristics compared to a community that is
rarely exposed to tropical cyclones. One interesting direction would be to
investigate if there is evidence of differences in probablity of disaster
declarations at geopolitical boundaries, like when comparing counties on either
side of a state border.

There are some cases where damage reports are directly used to infer the intensity 
of the weather event. For example, damage surveying is used following a tornado 
to assign the tornado's class on the [enhanced] Fujita scale. However, this is not the
intent of FEMA damage reports, which instead are used to determine post-disaster
aid [?] and are influenced by a number of political factors.

> "Because of the historic lack of direct measurements and remotely sensed tornado 
wind speeds at or near group level, damage surveying has remained the most common 
form for indicating tornado strength. ... The occurrence of a direct tornado strict 
upon a fixed, sufficiently sturdy, and well-calibrated wind measuring station is quite
rare. Only 31 direct in situ tornado observations are evident between 1894 and 2011."
[@edwards2013tornado]

@@

**Response:**

\begin{shaded}
\textbf{Reviewer 3 (R3) Comment 1:}
Starting on line 128, the authors write "Extreme winds are more common to the
track's right, where counterclockwise cyclonic winds move in concert with the
tropical cyclone's forward motion". I think this statement may be specific to
tropical cyclones in the Northern hemisphere (i.e., those relevant to exposure
in the US). Would be good to clarify this.
\end{shaded}

**Response:**

You are right, the opposite would be true in the Southern hemisphere. This
difference is driven by the Coriolis effect on cyclonic rotation, in which the
rotation of the earth diverts winds as they move toward the low pressure center
of the cyclone. We have edited the text to specify that this statement is
specific to the Northern Hemisphere:

> **Text in revised manuscript (relevant addition in bold):**
"**In the Northern Hemisphere, cyclonic winds are counterclockwise,** so
extreme winds are more common to the track's right where cyclonic winds move in
concert with the tropical cyclone's forward motion."

@@

> **Text in revised manuscript (relevant change in bold):**
"We used tracking data from HURDAT2, which records the storm center’s position
at **four standardized times for weather data collection (synoptic times),**
6:00 am, 12:00 pm, 6:00 pm, and 12:00 am Coordinated Universal Time (UTC)."

\begin{shaded}
\textbf{R3 Comment 3:}
Several of the exposures are calculated for the population mean center of each
county. From the reference list, it looks like the authors consistently used
population centers from 2010 throughout the years included in the dataset/R
package. If so, I suggest clarifying this in the text. 
\end{shaded}

**Response:**

This is correct---the population centers from the US 2010 Decennial Census were 
used throughout. We have added text to the manuscript to clarify this:

> **Text in revised manuscript (relevant addition in bold):**
"At each 15- interval, we measured the distance between the storm's
center and each county's population mean center, **as of the 2010 US Decennial
Census** (US Census Bureau 2020)."

@@

This database has evolved through several stages but has, since its earliest
iteration in the 1950s, served as the National Weather Service's [?] data of
record on tornado events. While there is also a database maintained by the 
[Severe Storm ...], the two databases have different names but almost 
identical information once data are aggregated to a daily, county-level 
measure of whether there was a tornado recorded in that county on that date. 
[Difference between the two and why there are two.]

There is no other tornado base at the same temporal and spatial scale that
we could use for validation.

This database serves as the basis for much of the large-scale (e.g., assessing
patterns over the US over long time periods) tornado-based research for the US. 
It does have some well-documented limitations. One is a notable increase over 
time in the number of tornadoes reported, particularly for the weakest, F0 [?]
tornadoes. [This stabilizes some in the age of radar?]

> "Weather events listed in the publication Storm Data are considered to be
the official database for tornadoes by the National Weather Service." 
[written by someone at the NWS Storm Prediction Center] [@mccarthy2003nws]

The US National Weather Service's Storm Prediction Center has a tornado database
called the "National Tornado Database" [@center2020storm]. This uses the NOAA
Storm Events database as its primary source [@center2020storm]. They note on their
website:

> "The tables below provide the links to comma separated values (.csv) files for
tornadoes, hail, and damaging winds, as compiled from *NWS Storm Data*. Tornado 
reports exist back to 1950 while hail and damaging wind events date from 1955.

The NOAA Storm Events database, or the Storm Prediction Center's version of the
tornado database have been used in a number of studies, including in a number of
studies of tornado climatology [@tippett2016more; @brooks2003climatological;
@strader2015climatology].

> "The dataset we will use is the so-called smooth log of severe-weather reports
collected by the SPC and archived in the National Oceanic and Atmospheric 
Administration publication *Storm Data*." [@brooks2003climatological]

> "Data such as tornado magnitude, path length, path width, and so on are recorded
by the National Climatic Data Center (NCDC) as a public service and for use
in meteorological, climatological, and engineering studies (Edwards et al. 2013)."
[@strader2015climatology]

Sometimes used in conjunction with other NWS data products:

> "Initially, tornado cases available in the NWS's Damage Assessment Toolkit
(NOAA, 2014) were employed, which is a GIS-based framework for collecting, storing, 
and retrieving damage survey data (Camp et al., 2014). This toolkit provides a 
variety of tornado event filtering (tornado survey point, track, footprint, and/or
swath) and download options [key mark-up language (.kml) or shapefile (.shp)], supplying
an initial sample of georeferenced damage assessments. In this study, a tornado 
footprint is defined as the maximum areal extent of tornado intensity inferred by 
the damage, wind speed measured directly by mobile Doppler radar, or assessed
theoretically as the recorded length multiplied with the maximum width as reported in 
Storm Data." [@strader2015climatology]

> "To serve the public interest and the National Climatic Data Center (NCDC) storm 
data record, the National Weather Service (NWS) documents the path length, width, 
and maximum damage rating for every tornado county segment (NOAA 2007).[Footnote
in manuscript: County segments of tornado paths are then combined at the NWS Storm
Prediction Center to yield a unified one-tornado (ONETOR) dataset of whole-tornado 
records (Schaefer and Edwards 1999).] Such data are used in meteorological and 
climatological research, as well as in determining construction standard for 
critical infrastructure such as high-tension electric lines and nuclear power plants
(e.g., Ramsdell et al. 2007)." [@edwards2013tornado]

Sometimes used in conjunction with other datasets:

> "The contiguous U.S. tornado fatality dataset utilized in this study was 
transcribed and compiled from two primary sources: (1) a long-term study of 
U.S. tornados by Grazulis (1993, 1997, hereafter Grazulis dataset) and 
(2) the National Climatic Data Center's Storm Data (NCDC 1959--2005) and 
'Storm Events' datavase. The Grazulis dataset included 'significant' tornado
events, including all events that are known to have produced a fatality, 
from 1680 to 1995. Storm Data reports from 1959 to 2005 were utilized to 
supplement the existing record of killer tornado events documented by 
Frazulis." [in a study of tornado fatalities] [@Ashley2007]

History:

> "During the early 1950s, the US Weather Bureau began a concerted effort to 
count all US tornadoes. In 1950, a list of tornadoes was provided in the 
*Climatological Data National Summary*, but it was not until 1953, the first 
full year of the issuance of Weather Bureau tornado watches, that the agency 
began to formally count all tornadoes. The *Climaticalogical Data National Summary*
document evolved into *Storm Data* by 1959, which, in addition to tornado 
data, included information on other storm perils, such as severe hail, high 
winds, and floods (Grazulis 1993). Over the years, the tornado dataset 
has been formatted and adjusted by the National Severe Storms Forecast 
Center and Storm Prediction Center (SPC), and currently the National Tornado 
Database is administered by the NWS Headquarters, SPC, and NCDC 
(McCarthy 2003)." [@Ashley2007]

@@

We have revised this figure to show an x-axis truncated at the minimum observed 
value ($\sim75\%$), as requested. The original and revised versions of this paper 
are reproduced in this response document as Figure \@ref(fig:fig3).

\begin{shaded}
\textbf{EHP Request 2:}
Main text figures (general)
To ensure that figures are accessible to readers with impaired color vision, do not use color as the sole means of conveying information in a figure unless absolutely necessary. Use different symbols, shading/textures, and line patterns instead of (or in addition to) color to distinguish among different data points. Use contrasting colors that can be easily distinguished from each other when the figure is printed in black and white. For details, see: \url{https://ehp.niehs.nih.gov/authors/figures}.
\end{shaded}

**Response:**

In all these figures, we have used colormaps that are both accessible to readers
with several forms of color vision deficiency, including the most common type
(deuteranomaly, a type of red-green color vision deficiency), and can be easily
distinguished when the figure is printed in black and white [@viridis;
@van2015mpl].

For these figures, we selected colormaps that were created to meet both these
criteria. These colormaps are all multi-hue sequential maps that avoid red-green
contrasts, while maximizing dynamic range in comparison to something
single-hued, like grayscale [@nunez2018optimizing; @viridis]. They are all
perceptually uniform across the scale for those with normal color vision
[@liu2018somewhere] and are close to perceptually uniform for those with common
forms of color deficiency [@nunez2018optimizing]. They reproduce their original
sequential gradient when printed in grayscale and so are perceptable when
printed on a black-and-white printer [@van2015mpl; @nunez2018optimizing].
Previous research has found that colormaps in this family allow viewers to more
quickly and accurately judge relative distances in scientific figures
[@liu2018somewhere]. The family of colormaps we used in the figures is currently
considered the gold standard for scientific figures [@nunez2018optimizing].

For example, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2) (left) of this
response document, we have reproduced the current version of Figure 5 as it
would appear to someone with three common types of color vision deficiencies:
deuteranomaly (defective green cone cells, the most common type), as well as
protanopia (defective red cone cells) and tritanopia (defective blue cone
cells). We have also shown how the Figure would appear in grayscale in
\@ref(fig:fig5check2) (right). We have included similar versions of other 
figures from the manuscript in response to later requests from EHP in this response.

```{r fig5check1, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", out.width="90%", fig.align = "center"}
include_graphics("figures/averageexposureonly_check1.pdf")
```

```{r fig5check2, fig.cap="Reproduction of revised version of Figure 5 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", out.width="90%", fig.align = "center"}
include_graphics("figures/averageexposureonly_check2.pdf")
```

@@

\begin{shaded}
\textbf{EHP Request 3:}
Figure 3: Please check the contrast between the colors used to indicate numbers
of counties. If you can see the gradient when printed in black and white it
should be ok, but some of the differences are very subtle. Truncating the x-axis
might help with this (if possible.)
\end{shaded}

**Response:**

As we note in our response to EHP Request 2, we used colormaps for all
figures that are designed to be perceptible both for those with common types of
color vision deficiency and when printed in grayscale. We have increased the
size of the circles in Figure 3 of the revised manuscript (reproduced in Figure
\@ref(fig:fig3) of this response) so that the colors are clearer, as well as
truncated the x-axis as requested in EHP Request 1. Also, we show in Figures
\@ref(fig:fig3check1) and \@ref(fig:fig3check2) of this response how the revised
Figure 3 would look with three common types of color vision deficiency as well
as when printed in black and white. The gradient is clear when this figure is
printed in black and white.

```{r fig3check1, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", out.width="100%"}
include_graphics("figures/windcomparison_check1.pdf")
```

```{r fig3check2, fig.cap="Reproduction of revised version of Figure 3 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", out.width="100%"}
include_graphics("figures/windcomparison_check2.pdf")
```

@@

**Response:**

\begin{shaded}
\textbf{EHP Request 5:}
Figures 5--6: Please check to see if the colors used can be distinguished if
printed in black and white, and if not, use a “color-blind friendly” palette if
possible. If this is not feasible, we can request and exemption from
accessibility requirements, but we try to do this only when alternatives  would
degrade the information content of  the figure.
\end{shaded}

**Response:**

As we note in our response to EHP Request 2, we used colormaps for all figures
that are designed to be perceptible both for those with common types of color
vision deficiency and when printed in grayscale Figures \@ref(fig:fig5check1)
and \@ref(fig:fig5check1) show what Figure 5 would look like under three common
types of color vision deficiency and when printed in black and white. Figures
\@ref(fig:fig6check1) and \@ref(fig:fig6check2) show the same thing for Figure
6. Therefore, these figures should not require any exemption from accessibility
requirements, as they should satisfy these requirements.

```{r fig6check1, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with two types of color vision deficiency, deuteranopia (left) and protanopia (right).", fig.show = "hold", out.width="50%"}
include_graphics("figures/jaccard_heatmap_check1.pdf")

@@

include_graphics("figures/jaccard_heatmap_check2.pdf")
```

```{r fig6check2, fig.cap="Reproduction of revised version of Figure 6 to check accessibility to readers with a third type of color vision deficiency, tritanopia (left) and for printing the figure in grayscale (right).", fig.show = "hold", out.width="50%"}
include_graphics("figures/jaccard_heatmap_check3.pdf")
include_graphics("figures/jaccard_heatmap_check4.pdf")
```
